shrimp_goby_boatnoise

### 2025-08-22 ms_boatnoise_shrimp_goby.Rmd
ensuring all formatting is unifrom

```{r}
# Generate data points
x <- seq(-20, 20, length.out = 1000)  # Adjusted range for better visualization

# Define parameters for the distributions
loc_logistic <- 0
scale_logistic <- 3

@@

# Calculate PDFs
pdf_logistic <- dlogis(x, location = loc_logistic, scale = scale_logistic)
pdf_student_t <- dt(x, df = df_student_t)
pdf_normal <- dnorm(x, mean = loc_normal, sd = scale_normal)

# Plot the PDFs
df <- data.frame(x = x, Logistic = pdf_logistic, Student_t = pdf_student_t, Normal = pdf_normal)
ggplot(df, aes(x)) +
  geom_line(aes(y = Logistic, color = "Logistic(0, 3)")) + # Default priors of Logistic(0,1) have been broadened as these priors seemed too restrictive
  geom_line(aes(y = Student_t, color = "Student_t(3, 0, 2.5)")) +
  geom_line(aes(y = Normal, color = "Normal(0, 3)")) + #For the beta coefficient prior, we will set it to be normally distributed around 0 with an sd of 5, here is the plotted prior distribution 
  labs(title = "Probability Density Functions",
       x = "x",
       y = "Probability Density") +
  scale_color_manual(name = "Distribution", values = c("Logistic(0, 3)" = "blue", "Student_t(3, 0, 2.5)" = "green", "Normal(0, 3)" = "red")) +
  theme_minimal()
```

@@

```{r}
#convergence diagnostics (R-hat)
extract_rhats <- function(model) {
  model_summary <- summary(model)
  #extract R-hat values for fixed effects
  fixed_rhats <- model_summary$fixed[, "Rhat", drop = FALSE]
  #check for random effects and extract R-hat values if present
  if (!is.null(model_summary$random)) {
    random_rhats <- do.call(rbind, lapply(model_summary$random, function(x) x[, "Rhat", drop = FALSE]))
    #combine fixed and random effects R-hat values
    rhats <- rbind(fixed_rhats, random_rhats)
  } else {
    rhats <- fixed_rhats
  }

@@

# Extract the relevant contrast information
goby_out_results.table <- as.data.frame(goby_out_contrast_effect) |> 
  dplyr::filter(contrast %in% c("control Pre - 2 Pre", "control Pre - 4 Pre","4 Pre - 2 Pre", "control Pre - control During", "control Pre - control Post", "4 Pre - 4 During", "4 Pre - 4 Post", "2 Pre - 2 During", "2 Pre - 2 Post"))

@@

#assigning a value for the baseline mean
goby_out_model_estimates_control_intercept <- goby_out_model_estimates_control |>
  dplyr::filter(stage == "Pre") |>
  pull(.value)

@@

```{r}
#posterior predictive checks
pp_check(shrimp_out_model)

@@

# Extract the relevant contrast information
shrimp_out_contrasts <- as.data.frame(shrimp_out_contrasts) |> 
  dplyr::filter(contrast %in% c("control Pre - 2 Pre", "control Pre - 4 Pre","4 Pre - 2 Pre", "control Pre - control During", "control Pre - control Post", "4 Pre - 4 During", "4 Pre - 4 Post", "2 Pre - 2 During", "2 Pre - 2 Post"))

@@

#assigning a value for the baseline mean
shrimp_out_model_estimates_control_intercept <- shrimp_out_model_estimates_control |>
  dplyr::filter(stage == "Pre") |>
  pull(.value)

@@

shrimp_contact_plot <- shrimp_contact_df |> 
  mutate(shrimp_contact = ifelse(is_zero, -0.1, shrimp_contact)) |> 
  ggplot(aes(x = shrimp_contact)) +
  geom_histogram(aes(fill = is_zero), binwidth = 0.05, 
                 boundary = 0, color = "white") +
  geom_vline(xintercept = 0)
shrimp_contact_plot
# Distribution is a zero-one inflated beta distribution, because the value for shrimp_contact is bounded by 0-1 as it is a proportion of the total time and there is values of 0 and 1.
```

@@

```{r}
#posterior predictive checks
pp_check(shrimp_contact_model)

@@

overall_mean <- shrimp_contact_model |>
  emmeans(~ 1,             # 1 means "no factors," i.e. the grand mean
          epred = TRUE,    # estimate on the response scale
          re_formla = NULL,
          rg.limit = 14000)

# View the summary with posterior mean and credible intervals
overall_mean

# Extract the relevant contrast information
shrimp_contact_contrasts <- as.data.frame(shrimp_contact_contrasts) |> 
  dplyr::filter(contrast %in% c("control Pre - 2 Pre", "control Pre - 4 Pre","4 Pre - 2 Pre", "control Pre - control During", "control Pre - control Post", "4 Pre - 4 During", "4 Pre - 4 Post", "2 Pre - 2 During", "2 Pre - 2 Post"))

@@

#Storing the model estimate means for each group to plot later on
shrimp_contact_model_estimates_control <- shrimp_contact_model_estimates |> 
  dplyr::filter(stroke == "control") |> 
  rename(.value = emmean)

@@

#assigning a value for the baseline mean
shrimp_contact_model_estimates_control_intercept <- shrimp_contact_model_estimates_control |>
  dplyr::filter(stage == "Pre") |>
  pull(.value)

@@

```{r}
#goby out
goby_out_contrasts <- goby_out_results.table |> 
  mutate(`Predictor variable` = "Treatment and phase interaction",
         contrast = str_replace_all(contrast, "\\b2\\b", "2-Stroke"),
         contrast = str_replace_all(contrast, "\\b4\\b", "4-Stroke"),
         contrast = str_replace_all(contrast, "\\bcontrol\\b", "Control")
         ) |> 
  rename(`Estimated marginal difference` = estimate,
         `Lower 95% CrI` = lower.HPD,
         `Upper 95% CrI` = upper.HPD,
         Contrast = contrast) |> 
  select(c(`Predictor variable`, Contrast, `Estimated marginal difference`, `Lower 95% CrI`, `Upper 95% CrI`))

@@

#shrimp out
shrimp_out_contrasts1 <- shrimp_out_contrasts |> 
  mutate(`Predictor variable` = "Treatment and phase interaction",
         contrast = str_replace_all(contrast, "\\b2\\b", "2-Stroke"),
         contrast = str_replace_all(contrast, "\\b4\\b", "4-Stroke"),
         contrast = str_replace_all(contrast, "\\bcontrol\\b", "Control")
         ) |> 
  rename(`Estimated marginal difference` = estimate,
         `Lower 95% CrI` = lower.HPD,
         `Upper 95% CrI` = upper.HPD,
         Contrast = contrast) |> 
  select(c(`Predictor variable`, Contrast, `Estimated marginal difference`, `Lower 95% CrI`, `Upper 95% CrI`))

@@

#shrimp contact
shrimp_contact_contrasts1 <- shrimp_contact_contrasts |> 
  mutate(`Predictor variable` = "Treatment and phase interaction",
         contrast = str_replace_all(contrast, "\\b2\\b", "2-Stroke"),
         contrast = str_replace_all(contrast, "\\b4\\b", "4-Stroke"),
         contrast = str_replace_all(contrast, "\\bcontrol\\b", "Control")
         ) |> 
  rename(`Estimated marginal difference` = estimate,
         `Lower 95% CrI` = lower.HPD,
         `Upper 95% CrI` = upper.HPD,
         Contrast = contrast) |> 
  select(c(`Predictor variable`, Contrast, `Estimated marginal difference`, `Lower 95% CrI`, `Upper 95% CrI`))

@@

# Extract the relevant contrast information
cumulative_sum_goby_out_results.table <- as.data.frame(cumulative_sum_goby_out_model_contrast_effect) |> 
  dplyr::filter(contrast %in% c("control Pre - 2 Pre", "control Pre - 4 Pre","4 Pre - 2 Pre", "control Pre - control During", "control Pre - control Post", "4 Pre - 4 During", "4 Pre - 4 Post", "2 Pre - 2 During", "2 Pre - 2 Post"))

@@

# Plot contrasts with HPD intervals
ggplot(goby_contrast_plot_data,
       aes(x = estimate, y = contrast, color = dataset)) +
  geom_point(position = position_dodge(width = 0.6), size = 2) +
  geom_errorbarh(aes(xmin = lower.HPD, xmax = upper.HPD),
                 position = position_dodge(width = 0.6), height = 0.2) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "black") +
  theme_classic() +
  labs(
    x = "Posterior median (with 95% HPD interval)",
    y = "Contrasts",
    title = "Comparison of contrasts between cumulative-sum and original models"
  )
```

@@

# Extract the relevant contrast information
cumulative_sum_shrimp_out_results.table <- as.data.frame(cumulative_sum_shrimp_out_model_contrast_effect) |> 
  dplyr::filter(contrast %in% c("control Pre - 2 Pre", "control Pre - 4 Pre","4 Pre - 2 Pre", "control Pre - control During", "control Pre - control Post", "4 Pre - 4 During", "4 Pre - 4 Post", "2 Pre - 2 During", "2 Pre - 2 Post"))

@@

# Plot contrasts with HPD intervals
ggplot(shrimp_contrast_plot_data,
       aes(x = estimate, y = contrast, color = dataset)) +
  geom_point(position = position_dodge(width = 0.6), size = 2) +
  geom_errorbarh(aes(xmin = lower.HPD, xmax = upper.HPD),
                 position = position_dodge(width = 0.6), height = 0.2) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "black") +
  theme_classic() +
  labs(
    x = "Posterior median (with 95% HPD interval)",
    y = "Contrasts",
    title = "Comparison of contrasts between cumulative-sum and original models"
  )
```

@@

# Extract the relevant contrast information
cumulative_sum_shrimp_contact_results.table <- as.data.frame(cumulative_sum_shrimp_contact_model_contrast_effect) |> 
  dplyr::filter(contrast %in% c("control Pre - 2 Pre", "control Pre - 4 Pre","4 Pre - 2 Pre", "control Pre - control During", "control Pre - control Post", "4 Pre - 4 During", "4 Pre - 4 Post", "2 Pre - 2 During", "2 Pre - 2 Post"))

@@

# Plot contrasts with HPD intervals
ggplot(shrimp_contact_contrast_plot_data,
       aes(x = estimate, y = contrast, color = dataset)) +
  geom_point(position = position_dodge(width = 0.6), size = 2) +
  geom_errorbarh(aes(xmin = lower.HPD, xmax = upper.HPD),
                 position = position_dodge(width = 0.6), height = 0.2) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "black") +
  theme_classic() +
  labs(
    x = "Posterior median (with 95% HPD interval)",
    y = "Contrasts",
    title = "Comparison of contrasts between cumulative-sum and original models"
  )
```

>>>>

```{r}
x <- seq(-20, 20, length.out = 1000)

loc_logistic <- 0
scale_logistic <- 3

@@

pdf_logistic <- dlogis(x, location = loc_logistic, scale = scale_logistic)
pdf_student_t <- dt(x, df = df_student_t)
pdf_normal <- dnorm(x, mean = loc_normal, sd = scale_normal)

df <- data.frame(x = x, Logistic = pdf_logistic, Student_t = pdf_student_t, Normal = pdf_normal)
ggplot(df, aes(x)) +
  geom_line(aes(y = Logistic, color = "Logistic(0, 3)")) + 
  geom_line(aes(y = Student_t, color = "Student_t(3, 0, 2.5)")) +
  geom_line(aes(y = Normal, color = "Normal(0, 3)")) +
  labs(title = "Probability Density Functions",
       x = "x",
       y = "Probability Density") +
  scale_color_manual(name = "Distribution", values = c("Logistic(0, 3)" = "blue", "Student_t(3, 0, 2.5)" = "green", "Normal(0, 3)" = "red")) +
  theme_minimal()
```

@@

```{r}
extract_rhats <- function(model) {
  model_summary <- summary(model)
  fixed_rhats <- model_summary$fixed[, "Rhat", drop = FALSE]
  if (!is.null(model_summary$random)) {
    random_rhats <- do.call(rbind, lapply(model_summary$random, function(x) x[, "Rhat", drop = FALSE]))
    rhats <- rbind(fixed_rhats, random_rhats)
  } else {
    rhats <- fixed_rhats
  }

@@

goby_out_results.table <- as.data.frame(goby_out_contrast_effect) |> 
  dplyr::filter(contrast %in% c("control Pre - 2 Pre", "control Pre - 4 Pre","4 Pre - 2 Pre", "control Pre - control During", "control Pre - control Post", "4 Pre - 4 During", "4 Pre - 4 Post", "2 Pre - 2 During", "2 Pre - 2 Post"))

@@

goby_out_model_estimates_control_intercept <- goby_out_model_estimates_control |>
  dplyr::filter(stage == "Pre") |>
  pull(.value)

@@

```{r}
pp_check(shrimp_out_model)

@@

shrimp_out_contrasts <- as.data.frame(shrimp_out_contrasts) |> 
  dplyr::filter(contrast %in% c("control Pre - 2 Pre", "control Pre - 4 Pre","4 Pre - 2 Pre", "control Pre - control During", "control Pre - control Post", "4 Pre - 4 During", "4 Pre - 4 Post", "2 Pre - 2 During", "2 Pre - 2 Post"))

@@

shrimp_out_model_estimates_control_intercept <- shrimp_out_model_estimates_control |>
  dplyr::filter(stage == "Pre") |>
  pull(.value)

@@

shrimp_contact_plot <- shrimp_contact_df |> 
  mutate(shrimp_contact = ifelse(is_zero, -0.1, shrimp_contact)) |> 
  ggplot(aes(x = shrimp_contact)) +
  geom_histogram(aes(fill = is_zero), binwidth = 0.05, 
                 boundary = 0, color = "white") +
  geom_vline(xintercept = 0)
shrimp_contact_plot
```

@@

```{r}
pp_check(shrimp_contact_model)

@@

overall_mean <- shrimp_contact_model |>
  emmeans(~ 1, epred = TRUE, re_formla = NULL, rg.limit = 14000)

overall_mean

shrimp_contact_contrasts <- as.data.frame(shrimp_contact_contrasts) |> 
  dplyr::filter(contrast %in% c("control Pre - 2 Pre", "control Pre - 4 Pre","4 Pre - 2 Pre", "control Pre - control During", "control Pre - control Post", "4 Pre - 4 During", "4 Pre - 4 Post", "2 Pre - 2 During", "2 Pre - 2 Post"))

@@

shrimp_contact_model_estimates_control <- shrimp_contact_model_estimates |> 
  dplyr::filter(stroke == "control") |> 
  rename(.value = emmean)

@@

shrimp_contact_model_estimates_control_intercept <- shrimp_contact_model_estimates_control |>
  dplyr::filter(stage == "Pre") |>
  pull(.value)

@@

```{r}
goby_out_contrasts <- goby_out_results.table |> 
  mutate(`Predictor variable` = "Treatment and phase interaction",
         contrast = str_replace_all(contrast, "\\b2\\b", "2-Stroke"),
         contrast = str_replace_all(contrast, "\\b4\\b", "4-Stroke"),
         contrast = str_replace_all(contrast, "\\bcontrol\\b", "Control")
         ) |> 
  rename(`Estimated marginal difference` = estimate,
         `Lower 95% CrI` = lower.HPD,
         `Upper 95% CrI` = upper.HPD,
         Contrast = contrast) |> 
  select(c(`Predictor variable`, Contrast, `Estimated marginal difference`, `Lower 95% CrI`, `Upper 95% CrI`))

@@

shrimp_out_contrasts1 <- shrimp_out_contrasts |> 
  mutate(`Predictor variable` = "Treatment and phase interaction",
         contrast = str_replace_all(contrast, "\\b2\\b", "2-Stroke"),
         contrast = str_replace_all(contrast, "\\b4\\b", "4-Stroke"),
         contrast = str_replace_all(contrast, "\\bcontrol\\b", "Control")
         ) |> 
  rename(`Estimated marginal difference` = estimate,
         `Lower 95% CrI` = lower.HPD,
         `Upper 95% CrI` = upper.HPD,
         Contrast = contrast) |> 
  select(c(`Predictor variable`, Contrast, `Estimated marginal difference`, `Lower 95% CrI`, `Upper 95% CrI`))

@@

shrimp_contact_contrasts1 <- shrimp_contact_contrasts |> 
  mutate(`Predictor variable` = "Treatment and phase interaction",
         contrast = str_replace_all(contrast, "\\b2\\b", "2-Stroke"),
         contrast = str_replace_all(contrast, "\\b4\\b", "4-Stroke"),
         contrast = str_replace_all(contrast, "\\bcontrol\\b", "Control")
         ) |> 
  rename(`Estimated marginal difference` = estimate,
         `Lower 95% CrI` = lower.HPD,
         `Upper 95% CrI` = upper.HPD,
         Contrast = contrast) |> 
  select(c(`Predictor variable`, Contrast, `Estimated marginal difference`, `Lower 95% CrI`, `Upper 95% CrI`))

@@

cumulative_sum_goby_out_results.table <- as.data.frame(cumulative_sum_goby_out_model_contrast_effect) |> 
  dplyr::filter(contrast %in% c("control Pre - 2 Pre", "control Pre - 4 Pre","4 Pre - 2 Pre", "control Pre - control During", "control Pre - control Post", "4 Pre - 4 During", "4 Pre - 4 Post", "2 Pre - 2 During", "2 Pre - 2 Post"))

@@

ggplot(goby_contrast_plot_data,
       aes(x = estimate, y = contrast, color = dataset)) +
  geom_point(position = position_dodge(width = 0.6), size = 2) +
  geom_errorbarh(aes(xmin = lower.HPD, xmax = upper.HPD),
                 position = position_dodge(width = 0.6), height = 0.2) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "black") +
  theme_classic() +
  labs(
    x = "Posterior median (with 95% HPD interval)",
    y = "Contrasts",
    title = "Comparison of contrasts between cumulative-sum and original models"
  )
```

@@

cumulative_sum_shrimp_out_results.table <- as.data.frame(cumulative_sum_shrimp_out_model_contrast_effect) |> 
  dplyr::filter(contrast %in% c("control Pre - 2 Pre", "control Pre - 4 Pre","4 Pre - 2 Pre", "control Pre - control During", "control Pre - control Post", "4 Pre - 4 During", "4 Pre - 4 Post", "2 Pre - 2 During", "2 Pre - 2 Post"))

@@

ggplot(shrimp_contrast_plot_data,
       aes(x = estimate, y = contrast, color = dataset)) +
  geom_point(position = position_dodge(width = 0.6), size = 2) +
  geom_errorbarh(aes(xmin = lower.HPD, xmax = upper.HPD),
                 position = position_dodge(width = 0.6), height = 0.2) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "black") +
  theme_classic() +
  labs(
    x = "Posterior median (with 95% HPD interval)",
    y = "Contrasts",
    title = "Comparison of contrasts between cumulative-sum and original models"
  )
```

@@

cumulative_sum_shrimp_contact_results.table <- as.data.frame(cumulative_sum_shrimp_contact_model_contrast_effect) |> 
  dplyr::filter(contrast %in% c("control Pre - 2 Pre", "control Pre - 4 Pre","4 Pre - 2 Pre", "control Pre - control During", "control Pre - control Post", "4 Pre - 4 During", "4 Pre - 4 Post", "2 Pre - 2 During", "2 Pre - 2 Post"))

@@

ggplot(shrimp_contact_contrast_plot_data,
       aes(x = estimate, y = contrast, color = dataset)) +
  geom_point(position = position_dodge(width = 0.6), size = 2) +
  geom_errorbarh(aes(xmin = lower.HPD, xmax = upper.HPD),
                 position = position_dodge(width = 0.6), height = 0.2) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "black") +
  theme_classic() +
  labs(
    x = "Posterior median (with 95% HPD interval)",
    y = "Contrasts",
    title = "Comparison of contrasts between cumulative-sum and original models"
  )
```
